Unified data structures to optimize solving a complex of interrelated geometric problems*\\

### 2019-11-17 article/article.tex
updated introduction 3 (by VT)

The notion of a convex hull is simple. For a set of points $S$ in a $k$-dimensional space it is a smallest convex set, that comprise $S$. The problem of building such a set has a long history. In practice to solve such problem, means to find a subset in $S$, which can be a "skeleton" for the convex hull.

It is differentiated between two types of point sets in this case: a static and a dynamic one. In the first case all points are known in advance. The algorithms for solving that kind of problem are well researched and described in \cite{jarvis, graham, quickhull}. In the letter case an algorithm should be able to handle new points which are being added to the initial set and some  being removed from it in real time. \textcolor{teal}{Initially a set of points can either contain points or be empty.}

The first known algorithm for solve such problem was introduces by Preparata in \cite{preparata}. Convex hull is maintained in a form of an AVL tree, that allows to perform operations of adding new points form it. The split and finding of tangents is performed in optimal time of $O(\log n)$. The major drawback of this method is that there is no mean to remote points from the set.

Soon after Overmars and van Leeuwen described a method of how a concatenable queue can be used to maintain the convex hull \cite{overmars}. This, as well as special data structure that is based on a balanced binary tree and maintains the information gathered during previous queries, allows to support fully dynamic sets of points and perform add or remove query  $O(\log^2 n)$ in the worst case.

% new algorithm
This article introduces a new approach for solving the convex hull problem on a static set of points. Here the ``divide-and-conquer'' strategy is used and the actual convex hull is maintained in a form of a concatenable queue during the computations. The resulting algorithm has the worst case complexity of $O(n\log n)$, whereas the subroutine of merging 2 convex hull has a complexity of $O(log n)$, where $n$ denotes the number of points.

% unified algorithmic environment
\textcolor{teal}{The convex hull problem arises not only in many practical applications, but also is an core method for solving many others computational geometry problems. There are implicit relations between such problems.} Currently, the state of the art approach for solving such sets of problems is the single algorithmic environment, which takes advantage of those implicit relations to implement efficient solving methods. An example to such approach may be the usage of a generalized data structures that can be used in different algorithms. \textcolor{teal}{Thus, the algorithmic environment allows to reduce complex problems to some basic computations.}

In the algorithmic environment the ``divide-and-conquer'' algorithm is formalizes from a programmatic standpoint. The formalization allows to create a level of abstraction upon an actual implementation of an algorithm and therefore perform computations sequentially and in parallel.

\section{Convex hull algorithm}
\subsection{Previous work}

>>>>

Nowadays, advanced computer simulations and visualizing of complex scientific researches as well as large scale technical projects requires to simultaneously solve a set of \textcolor{red}{science and technology objectives}. The core of this set are problems of computational geometry and computer graphics. To solve such problems it is needed to create suitable algorithmic frameworks, that would yield accurate results in real time. Existing methods (iLastic[1], IMARIS [2], ImageJ (або FiJi) [3]), that are based on set of algorithms implementations organized in a package does not result in desirable efficiency and accuracy. It is worse noting, that there are a lot of parallel algorithms designed to solve specifically certain computational geometry problems [4-14]. Every such algorithm requires its own computational resources and is executed independently from others. In such case some identical steps, such as precomputation and building data structures, are performed several times.

Therefore, a important objective in developing of the algorithmic models is to create a universal tool, which would have means to efficiently solve a problems in the set. This tool should also execute identical steps of the algorithms once and be able to represent results of those steps in a form of the common data structures. In [15] the notion of a single algorithmic environment is introduced, which is based on the ``divide-and-conquer'' principle and takes into account the aforementioned features of the algorithms. In particular, the precomputation and splitting the initial set of data to form the recursion tree is common for all the problems is as s result is executed only once. During the merge stage the intermediate results are maintained in the weighted concatenable queue. This model does not repeat computations and the intermediate results are highly reused during the algorithms, which yields good performance.

\textcolor{blue}{This article provides detailed explanation on how the common data structures from [15] are implemented on the example of a convex hull algorithm.}

\section{Convex hull algorithm}
\subsection{Problem statement}

The notion of a convex hull is simple. For a set of points $S$ in a $k$-dimensional space it is a smallest convex set, that comprise $S$. In practice to solve such problem, means to find a subset in $S$, which can be a "skeleton" for the convex hull.
